---
editor: 
  markdown: 
    wrap: 72
---

# Cleans your data per your specifications and converts to tidy format

Joining since we did not need to group together two dataset. We won't
use the joined dataset for the rest of the project,

it is just a demonstration of our joining skills

## Addition of an observation ID

```{r}
dat_load_01 <- dat_load_01 |> 
  mutate(ID = row_number()) |> 
  relocate(ID)
  
```

## Separation of the dataset in two distinct datasets

```{r}
df_Etiology <- dat_load_01 |> 
  select(ID,
         Alcohol,
         hcv,
         hbv,
         nafld,
         Cryptogenic,
         PrimarySclerosingCholangitis,
         Autoimmunehepatitis,
         pbc,
         Otherliverdiseaseetiology
         )

df_Comorbidities <- dat_load_01 |> 
  select(ID,
         mi,
         chf,
         pvd,
         Cerebrovasculardisease,
         Dementia,
         Chronicpulmonorydisease,
         Rheumatologicaldisease,
         Pepticulcerdisease,
         Anymalignancyincludinglymphomaor,
         Moderateorsevereliverdisease,
         aids
         )
```

## Joining the two datasets

```{r}
df_Etiology_and_Comorbidities <- left_join(df_Etiology, df_Comorbidities)
df_Etiology_and_Comorbidities
```

## Tidying of the data using pivot_longer

```{r}

dat_clean_02 <- dat_load_01 |> 
  # Selection of correct variable 
  select(Sex, 
         `Age(Decade)atadmission`,
         Liverrelateddiagnosis1, 
         Alcohol, 
         mi,
         chf, 
         pvd, 
         Cerebrovasculardisease,
         Dementia, 
         Chronicpulmonorydisease, 
         Rheumatologicaldisease,
         Pepticulcerdisease,
         Mildliverdisease,
         Diabetes,
         Diabeteswithchroniccomplications,
         Heiplegiaorparaplegia,
         Renaldisease,
         Anymalignancyincludinglymphomaor, 
         Moderateorsevereliverdisease,
         Metastaticsolidtumor,
         aids, 
         cps, 
         Mortality_InHospital, 
         CHARLSON_CI,
         MORTALITY_30_DAYS_postDC, 
         MORTALITY_90_DAYS_postDC ) |> 
         
  mutate(C_Diabetes_with_without_complications = Diabetes + Diabeteswithchroniccomplications) |> 
  select(-c(Diabetes, Diabeteswithchroniccomplications)) |> 
  
  # Change the name of variables 
  rename(Age = `Age(Decade)atadmission`, 
         Liver_disease = Liverrelateddiagnosis1, 
         Alcohol_consumption = Alcohol, 
         C_Myocardial_infarction = mi, 
         C_Congestive_Heart_Failure = chf, 
         C_Peripheral_vascular_disease = pvd, 
         C_Cerebro_Vascular_disease = Cerebrovasculardisease, 
         C_Dementia = Dementia, 
         C_Chronic_pulmonary_disease = Chronicpulmonorydisease, 
         C_Rheumatological_disease = Rheumatologicaldisease,
         C_Pepticulcer_disease = Pepticulcerdisease, 
         C_Mild_liver_disease = Mildliverdisease, 
         C_Heiplegia_paraplegia = Heiplegiaorparaplegia, 
         C_Renal_disease = Renaldisease, 
         C_Malignancy = Anymalignancyincludinglymphomaor, 
         C_Moderate_severe_liver_disease = Moderateorsevereliverdisease, 
         C_Metastatic_solid_tumor = Metastaticsolidtumor, 
         C_AIDS = aids, 
         CPS = cps, 
         Mortality_hospital = Mortality_InHospital, 
         Charlson_CI = CHARLSON_CI, 
         Mortality_30_days = MORTALITY_30_DAYS_postDC, 
         Mortality_90_days = MORTALITY_90_DAYS_postDC, 
         ) |> 

  # Reordering columns 
  relocate(C_Diabetes_with_without_complications, .after = C_AIDS) |> 
  relocate(Charlson_CI, .after = CPS) |> 

  # Add a column "C_none"
  mutate(C_None = (C_AIDS == 0) &
        (C_Cerebro_Vascular_disease ==0)&
        (C_Congestive_Heart_Failure == 0) &
        (C_Chronic_pulmonary_disease == 0) & 
        (C_Dementia == 0) &
        (C_Diabetes_with_without_complications == 0) & 
        (C_Heiplegia_paraplegia == 0) & 
        (C_Malignancy == 0) &
        (C_Myocardial_infarction == 0) &
        (C_Mild_liver_disease == 0) & 
        (C_Moderate_severe_liver_disease == 0) & 
        (C_Metastatic_solid_tumor == 0) & 
        (C_Pepticulcer_disease == 0) & 
        (C_Peripheral_vascular_disease == 0) & 
        (C_Rheumatological_disease == 0) & 
        (C_Renal_disease == 0)
      ) |> 


  # Change 1/0 into Yes/No
  mutate(Alcohol_consumption = ifelse(Alcohol_consumption == 1, 'Yes', 'No')) |> 
  mutate_at(vars(starts_with('C_')), ~ifelse(. == 1, 'Yes', 'No')) |> 
  mutate_at(vars(starts_with('Mortality')), ~ifelse(. == 1, 'Yes', 'No')) |> 


  # If mortalityH = yes -> mortality 30-90 = no 
  mutate_at(vars(c(Mortality_30_days, Mortality_90_days)), ~ifelse(Mortality_hospital == 'Yes', 'No', .)) |> 
  #If mortality30 = yes -> mortality 90 = no   
  mutate_at(vars(Mortality_90_days), ~ifelse(Mortality_30_days == 'Yes', 'No', .)) |>
  
  # Create a id for each ind 
  mutate(Id = row_number()) |>
  relocate(Id, .before = Sex) |> 
 
  # Create a column Comorbidities ou pivot longer ? definition de tidy
  pivot_longer(
    cols = starts_with("C_"), 
    names_to = "Comorbidities", 
    values_to = "Presence_of_comorbidities",
    values_drop_na = TRUE) |> 
  
  filter(Presence_of_comorbidities == 'Yes') |> 
  select(-Presence_of_comorbidities) |> 
  drop_na(CPS) |> 
  drop_na(Sex) |> 
  mutate(Comorbidities = str_remove(Comorbidities, "C_"))

```

```{r}
dat_clean_02
```


## Save data in the folder "data"

```{r}
write_tsv(
  x= dat_clean_02, 
  file = '~/projects/Group_19_project/data/02_dat_clean.tsv'
  )
```
