---
title: "03_augment"
format: html
editor: visual
---

## Introduction of a new variable "Mortality"


We have two data sets available: a clean and a tidy one. We will augment both of
them simultaneously.
We need to be careful since there is an 'error' in the database where people 
that died at time t are written as dead at t + 1. However, we want the mortality
to assess the exact mortality time.

The combination are as follows:  

    H No 30 No 90 No = "Alive", 
    H Yes 30 No 90 No = "Dead_Hospital" 
    H Yes 30 Yes 90 Yes = "Dead_Hospital" 
    H No 30 No 90 Yes = "Dead_90" 
    H No 30 Yes 90 Yes = "Dead_30"

### Clean data set
```{r}

dat_aug_clean_03 <- dat_clean_02 |> 
   mutate(Mortality = case_when(Mortality_hospital == 'Yes' ~ 
                                "Death_in_hospital", 
                                Mortality_hospital == 'No' & 
                                Mortality_30_days == 'Yes' & 
                                Mortality_90_days == 'No' ~ 
                                "Death_in_first_30_days", 
                                Mortality_hospital == 'No' & 
                                Mortality_30_days == 'No' & 
                                Mortality_90_days == 'Yes' ~ 
                                "Death_between_30_and_90_days", 
                               .default = "No death")) |> 
   select(-c(Mortality_30_days, Mortality_90_days, Mortality_hospital))
dat_aug_clean_03
```


### Tidy data set
```{r}
dat_aug_03 <- dat_tidy_02 |> 
   mutate(Mortality = case_when(Mortality_hospital == 'Yes' ~ "Death_in_hospital", 
                                Mortality_hospital == 'No' & Mortality_30_days == 'Yes' & Mortality_90_days == 'No' ~ "Death_in_first_30_days", 
                                Mortality_hospital == 'No' & Mortality_30_days == 'No' & Mortality_90_days == 'Yes' ~ "Death_between_30_and_90_days", 
                               .default = "No death")) |> 
   select(-c(Mortality_30_days, Mortality_90_days, Mortality_hospital))
dat_aug_03
```

## Introduction of a categorical variable - change Decades into age categories

### Clean
```{r}
dat_aug_clean_03 <- dat_aug_clean_03 |>
  mutate(Age = str_c('[', Age*10, '; ', (Age + 1)*10, ']'))

dat_aug_clean_03
```

### Tidy
```{r}
dat_aug_03 <- dat_aug_03 |>
  mutate(Age = str_c('[', Age*10, '; ', (Age + 1)*10, ']'))

dat_aug_03
```

## Save data in the folder "data"

```{r}
write_tsv(
  x= dat_aug_clean_03, 
  file = '~/projects/Group_19_project/data/03_dat_aug_clean.tsv'
  )
```

```{r}
write_tsv(
  x= dat_aug_03, 
  file = '~/projects/Group_19_project/data/03_dat_aug.tsv'
  )
```

