---
title: "05_analysis_mortality"
format: html
editor: visual
---

# Effect of different variables on mortality

During that part, we will investigate to determine to what extent some variables
like CPS, age and liver disease affect the mortality of the patients before 
investigating specific and relevant etiologies that have been picked using a 
linear regression performed in file 07_analysis_CPS.qmd.

## CPS Mortality

```{r}
bar_plot_CPS_Mortality<- dat_aug_03 |> 
  ggplot(aes(x = Mortality,
             y = CPS,
             fill = Mortality)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Boxplot chart of CPS values",
       subtitle = "stratified by mortality",
       y = "CPS") +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank()) +
  scale_x_discrete(guide = guide_axis(angle = 25))

bar_plot_CPS_Mortality

```

This plot proves the fact that CPS value accurately describes the mortality of 
cirrhosis patients. It shows that a low CPS is correlated to a higher survival 
rate. On the other side, a patient with a high CPS has more chances of dying of 
cirrhosis.

### Saving the plot

```{r}
ggsave("CPS_and_mortality_boxplot.png", 
       path = "../results/plots/",
       plot = bar_plot_CPS_Mortality,
       width = 7,
       height = 5)
```

## Age impact on mortality

First, we will see how the age of the patients impact their mortality by 
plotting the distribution of patient outcomes stratified by age.

```{r}
Age_impact_mortality1 <- dat_aug_03 |> 
  ggplot(mapping = aes(x = Age,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "stack") +
  scale_fill_viridis_d() +
  labs(title = "Distribution of patient outcomes ",
       subtitle = "stratified by age",
       x = "Age",
       y = "Count") +
  theme_minimal()

Age_impact_mortality1
```

```{r}
Age_impact_mortality2 <- dat_aug_03 |> 
  ggplot(mapping = aes(x = Age,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Proportions of patient outcomes ",
       subtitle = "stratified by age",
       x = "Age",
       y = "Proportions"
)

Age_impact_mortality2
```

The plot showing the proportions of patient outcomes is more relevant than the 
one showing their distribution, because it is relative to the count of patients.
Nonetheless, the distribution plot allows us to see that the age categories 
\[30,40\] and \[90,100\] don't contain enough patients to be statistically 
relevant. Therefore, if we only take the other age categories into 
consideration, we can see slight increase in mortality when the age of the 
patients is increasing.


### Saving the plots
```{r}
ggsave("Age_Mortality_count.png", 

       path = "../results/plots/",
       plot = Age_impact_mortality1,
       width = 7,
       height = 5)

ggsave("Age_Mortality_prop.png", 
       path = "../results/plots/",
       plot = Age_impact_mortality2,
       width = 7,
       height = 5)

```

## Effect of different liver diseases on mortality

```{r}
bar_plot_LD_Mortality<- dat_aug_03 |> 
  ggplot(aes(x = Liver_disease, 
             fill = Mortality)) +
  geom_bar(position = "fill", 
           stat = "count", 
           alpha = 0.7) +
  labs(title = "Liver disease repartition in cirrhosis patients, stratified 
       by mortality",
       y = "Proportion") +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.title.x=element_blank()) +
  scale_x_discrete(guide = guide_axis(angle = 25))

bar_plot_LD_Mortality
```

By looking at the mortality of patients with different liver diseases, we can 
see the type of liver disease impacts greatly the mortality of patients. The 
size of the samples "Hepato-hydrothorax" and "Jaundice- as isolated 
presentation" being small (less than 15 patients), the data concerning these 
liver disease should not be taken into account for the analysis. This plot also
provides some information about how fulminant the different liver diseases are. 
For instance, the categories "Other fluid overload" and "SBP" have approximately
the same proportions of overall mortality, but the category "Other fluid 
overload" seems to be more fulminant because all of the patients that died after
contracting that disease died in hospital, when the majority of the patient that
died after contracting SBP died more than 30 days after leaving the hospital.

### Saving the plot

```{r}
ggsave("LD_Mortality.png", 
       path = "../results/plots/",
       plot = bar_plot_LD_Mortality,
       width = 7,
       height = 5)
```


# Effect of different etiologies on mortality

We start by creating a new data set where we add three variables. The three 
variable are three different etiologies and their significance concerning Child 
Pugh score was assessed with a linear regression model in the file 07. Because 
the Child Pugh Score indicates the severity of the cirrhosis, it could be 
related to the mortality, and we will verify that hypothesis with the following
plots.

```{r}
dat_mortality <- dat_aug_03 |> 
  mutate(Alcohol = ifelse(Etiologies == "Alcohol_consumption", "yes", "no")) |> 
  mutate(NAFLD = ifelse(Etiologies == "Non_alcoolic_fatty_LD", "yes", "no")) |> 
  mutate(Cryptogenic = ifelse(Etiologies == "Cryptogenic", "yes", "no")) 

dat_mortality

```

## Alcohol consumption impact on mortality

To determine if alcohol consumption impacts mortality, we plot the distribution 
and the proportions of patient outcomes stratified by alcohol consumption. 
Alcohol was considered by the linear regression as a non-significant etiology, 
we therefore expect to see very few changes in the proportions. We will here 
plot the distribution (with normal counts) and the proportion of the etiology in
the population to be able to have a relative comparison. 

```{r}
Alcohol_distribution <- dat_mortality |> 
  ggplot(mapping = aes(x = Alcohol,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "stack") +
  scale_fill_viridis_d() +
  labs(title = "Patient outcomes",
       subtitle = "stratified by alcohol consumption",
       x = "Alcohol consumption", 
       y = "Count") + 
  theme(legend.position = "none")


Alcohol_proportions <- dat_mortality |> 
  ggplot(mapping = aes(x = Alcohol,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "fill") +
  scale_fill_viridis_d() +
  labs(
    title = "Proportions of patient outcomes ",
    subtitle = "stratified by alcohol consumption",
    x = "Alcohol consumption",
    y = "Proportions")

Alcohol_impact_mortality <- ggarrange(Alcohol_distribution,
                                      Alcohol_proportions, 
                                      legend = 'bottom',
                                      common.legend = TRUE)
                                
Alcohol_impact_mortality 
```

The proportions plot shows a very slight/ unsignificant increase in mortality 
when the patients are alcohol consumers, therefore we can't notice a big
mortality difference between the two populations. 

### Saving the plot

```{r}
ggsave("Alcohol_impact_mortality.png", 
       path = "../results/plots/",
       plot = Alcohol_impact_mortality,
       width = 7,
       height = 5)
```

## Non-alcoholic fatty liver disease impact on mortality

On the contrary to alcohol consumption, NAFLD was considered a significant 
etiology. 

```{r}
NAFLD_distribution <- dat_mortality |> 
  ggplot(mapping = aes(x = NAFLD,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "stack") +
  scale_fill_viridis_d() +
  labs(title = "Patient outcomes",
       subtitle = "stratified by presence of NAFLD",
       x = "Patient suffering from NAFLD", 
       y = "Count") + 
  theme(legend.position = "none")

NAFLD_proportions <- dat_mortality |> 
  ggplot(mapping = aes(x = NAFLD,
                       fill = Mortality,
                       na.rm = TRUE)) +
  geom_bar(alpha = 0.6,
           position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Proportions of patient outcomes",
       subtitle = "stratified by presence of NAFLD",
       x = "Patient suffering from NAFLD",
       y = "Proportions"
)

NAFLD_impact_mortality <- ggarrange(NAFLD_distribution,
                                      NAFLD_proportions, 
                                      legend = 'bottom',
                                      common.legend = TRUE)
NAFLD_impact_mortality
```
 
On the proportions plot, we can see that patients with NAFLD seem to have more 
chance to survive to cirrhosis. However, we can't reach any specific conclusion 
due to the small population of patients with NAFLD.
NAFLD, significant etiology, has therefore a 'positive' impact on cirrhosis 
patients, which seems hard to understand.

### Saving the plot

```{r}
ggsave("NAFLD_impact_mortality.png", 
       path = "../results/plots/",
       plot = NAFLD_impact_mortality,
       width = 7,
       height = 5)
```

## Conclusion 

This analysis allows to conclude that CPS and the type of liver disease have a
major impact on patient's outcome, whereas age has a slight impact. We also got
to test the linear regression significant results, noticing that NAFLD seems to
have a positive impact on the mortality, whereas alcohol consumption doesn't 
seem to play a role in the patients outcome.
